Parsing Postfix log files

### 2007-06-24 doc/logparser.tex
Wrote most of the rule efficiency section.  It needs a bit more work and proof reading, but it's mostly there now.  Changed the graphs slightly, added new graphs, and updated the Makefile.

% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}
\renewcommand{\refname}{Bibliography}
% Add the bibliography into the table of contents.
\usepackage[section,numbib,nottoc]{tocbibind}
% When creating a PDF make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.
\usepackage{hyperref}
\hypersetup{%
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final,
    pdfborder   = {0 0 0},
    breaklinks,
}

@@

XXX LOOK AT citeseer.com AND scholar.google.com FOR RELATED PAPERS\@.
REVIEW EACH PAPER, WRITING AN ABSTRACT FOR IT\@: WHAT'S THE SAME, WHAT'S
DIFFERENT, PERFORMANCE AND \newline COMPLETENESS COMPARISON\@.  FROM THIS
I'LL GET AN OVERALL PICTURE, AND I CAN WORK REFERENCES TO THE PAPERS INTO
THIS PAPER, PLUS MAYBE A PRIOR ART SECTION\@.

@@

@@

    \item [rule\_order] is an efficiency measure.  This counter is
        maintained for every rule and incremented each time the rule
        successfully matches.  At the start of each run the program sorts
        the rules in descending rule\_order, and at the end of the run
        updates every rule's rule\_order.  Assuming that the distribution
        of log lines is reasonably consistent between log files, rules
        matching more commonly occurring log lines will be tried before
        rules matching less commonly occurring log lines, lowering the
        program's execution time.  Rule efficiency concerns are discussed
        in section~\ref{rule efficiency}.

@@

@@

XXX WRITE ABOUT\@: DYNAMICS OF THE SYSTEM

XXX PRIORITY, OVERLAPPING RULES, MORE SPECIFIC VS \newline MORE GENERAL

XXX EXPLAIN THE DIPS IN THE GRAPHS WITH LARGE LOG FILES

XXX MENTION THAT THE PROGRAM HAS LINER COMPLEXITY AND SCALES LINEARLY WITH
THE LOG FILE SIZE

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 50 MB log files containing 300,000 log lines.  The graphs in
section~\ref{graphs} illustrate the execution time for various
configurations of the parser and should be referred to during the following
discussion.

An important property of a parser is how execution time scales with input
size: does it scale linearly, polynomially, or exponentially?
Graph~\ref{execution time vs file size vs number lines} shows the execution
time in seconds, file size in MB and tens of thousands of lines per log
file.  All three lines run roughly parallel, giving a visual impression
that the algorithm scales linearly with input size.  This is borne out by
graph~\ref{execution time vs file size vs number lines factor}, which plots
the ratio of file size in MB vs execution time in seconds and number of
lines

Graph~\ref{normal regex vs discard regex} shows execution times gathered
over 93 log files, with and without caching the regex; also shown on the
graph are the file sizes and number of lines in each file.
Graph~\ref{normal regex vs disarded regex factor} shows the percentage
execution time increase when not caching each regex.  Each log file was
parsed 10 times, the first run discarded and the remaining 9 runs averaged.
Logging to the database was disabled for the test runs.

Graph~\ref{normal vs shuffled vs reversed ordering} plots the execution
time over 93 log files for normal, shuffle and reverse ordering.
Graph~\ref{normal vs shuffled vs reversed ordering factor} shows the
percentage execution time increase for shuffled and reversed over normal.
Each log file was parsed 10 times, the first run discarded and the
remaining 9 runs averaged.  Logging to the database was disabled for the
test runs.

It is hoped that this will prove to have greater effectiveness as the
number of rules grows; currently the effect is negligible in comparison to
caching each compiled regex, but it still provides approximately a 7--10\%
performance increase.

@@

\bibliographystyle{logparser-bibliography-style}

XXX ADD RESOURCES/REFERENCES HERE AND THEN MAKE REFERENCES TO THEM FROM WITHIN THE PAPER

\label{bibliography}
\bibliography{logparser-bibliography}

@@

\subsection{Execution time vs file size vs number of lines.}
\label{execution time vs file size vs number lines}
\includegraphics{plot-normal-filesize-numlines.ps}

\subsection{Ratio of file size and number of lines to execution time}
\label{execution time vs file size vs number lines factor}
\includegraphics{plot-normal-filesize-numlines-factor.ps}

\subsection{Regex: cached vs discarded}
\label{normal regex vs discard regex}
\includegraphics{plot-cached-discarded.ps}

\subsection{Regex: percentage execution time increase}
\label{normal regex vs disarded regex factor}
\includegraphics{plot-cached-discarded-factor.ps}

\subsection{Normal vs shuffled vs reversed ordering}
\label{normal vs shuffled vs reversed ordering}
\includegraphics{plot-normal-shuffle-reverse.ps}

\subsection{Percentage increase of shuffled and reversed over normal}
\label{normal vs shuffled vs reversed ordering factor}
\includegraphics{plot-normal-shuffle-reverse-factor.ps}

>>>>

% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}
\renewcommand{\refname}{Bibliography}
% Add the bibliography into the table of contents.
\usepackage[section,numbib,nottoc]{tocbibind}
% When creating a PDF make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.
\usepackage{hyperref}
\hypersetup{%
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0 0 0},
    breaklinks  = true,
}

@@

XXX MAYBE A PRIOR ART SECTION\@?

@@

XXX MAYBE ADD A SECTION ABOUT OPTIMISATION OF THE DATABASE\@.

@@

    \item [rule\_order] is an efficiency measure.  This counter is
        maintained for every rule and incremented each time the rule
        successfully matches.  At the start of each run the program sorts
        the rules in descending rule\_order, and at the end of the run
        updates every rule's rule\_order.  Assuming that the distribution
        of log lines is reasonably consistent between log files, rules
        matching more commonly occurring log lines will be tried before
        rules matching less commonly occurring log lines, lowering the
        program's execution time.  Rule efficiency concerns are discussed
        in section~\ref{rule ordering for efficiency}.

@@

XXX WRITE ABOUT DATABASE OPTIMISATION??

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 50 MB log files containing 300,000 log lines (for a small mail
server --- large scale mail servers would have much larger log files on a
daily basis).  When generating the data for the graphs included in
section~\ref{graphs}, 93 log files (totaling 10.08 GB, 60.72 million lines)
were each parsed 10 times, the first run discarded and the remaining 9 runs
averaged.  The first run is discarded for two reasons:

    \item The execution time will be higher because the log file must be
        read from disk, whereas for subsequent runs the log file will be
        cached in memory by the operating system.

    \item The execution time will also be higher because the rule ordering
        will be sub-optimal compared to subsequent runs.

Saving results to the database was disabled for the test runs, as that
dominates the run time of the program, and the tests are aimed at measuring
the speed of the parser rather than the speed of the database.

\subsubsection{Algorithmic complexity}

An important property of a parser is how execution time scales with input
size: does it scale linearly, polynomially, or exponentially?
Graph~\ref{execution time vs file size vs number lines} shows the execution
time in seconds, file size in MB and tens of thousands of lines per log
file per log file.  All three lines run roughly in parallel, giving a
visual impression that the algorithm scales linearly with input size.  This
impression is borne out by graph~\ref{execution time vs file size vs number
lines factor}, which plots the ratio of file size vs execution time and
number of lines vs execution time.  As the reader can see, the ratios are
quite tightly banded, showing that the algorithm scales linearly: the much
larger log files between points 60 and 70 on the X axis in
graph~\ref{execution time vs file size vs number lines} don't cause any
abnormality in the corresponding points in graph~\ref{execution time vs
file size vs number lines factor}.

\subsubsection{Rule ordering for efficiency}

\label{rule ordering for efficiency}

XXX DO A FULL RUN ONCE RULE\_ORDER ISN'T BEING TRASHED AND GRAPH RULE
HITS\@.

Rule ordering was briefly discussed in section~\ref{rule attributes} and
it will be covered in greater detail in this section.  At the time of
writing there are 131 different rules, some of which match much more
frequently than others.  Assuming that the distribution of log lines is
reasonably steady over time, program efficiency should benefit from trying
more frequently matching rules before those which match less frequently.
To test this hypothesis three full test runs were performed with different
rule orderings:

    \item [normal]  The most optimal order, according to the hypothesis:
        rules which match most often will be tried first.

    \item [shuffle] Random ordering --- the rules will be shuffled once
        before use and will retain that ordering for the entirety of the
        log file.  Note that the ordering will change every time the parser
        is executed, so 10 different orderings will be generated for each
        log file in the test run.  This is intended to represent an
        unsorted rule set.

    \item [reverse] Hypothetically the least optimal order: the most
        frequently matching rules will be tried last.

The resulting execution times are plotted in graph~\ref{normal vs shuffled
vs reversed ordering}, where the difference is not readily apparent, but
graph~\ref{normal vs shuffled vs reversed ordering factor} shows the
percentage increase of execution times:

    \item [shuffled] 6--11\% increase; mean 8.34, standard deviation 2.44.

    \item [reversed] 15--22\% increase; mean 17.97, standard deviation
        2.94.

Overall this increases the parser's efficiency by 5.5--10\%; a modest but
worthwhile increase.

\subsubsection{Caching each regex}

Graph~\ref{normal regex vs discard regex} shows execution times with and
without caching the regex; also shown on the graph are the file sizes and
number of lines in each file.  Caching the compiled regexs is obviously far
more efficient; graph~\ref{normal regex vs disarded regex factor} shows the
percentage execution time increase when not caching each regex is between
350--570\% (mean 465.58\%, standard deviation 114.70\%).  Perl compiles the
original regex into an internal representation, optimising the regex to
improve the speed of matching, but this compilation and optimisation takes
CPU time.  Perl automatically caches static regexs, but dynamic regexs need
to be explicitly compiled and cached.

@@

\bibliographystyle{logparser-bibliography-style}
\label{bibliography}
\bibliography{logparser-bibliography}

@@

\subsection{Parser scalability}

\subsubsection{Execution time vs file size vs number of lines.}
\label{execution time vs file size vs number lines}
\includegraphics{plot-normal-filesize-numlines.ps}

\subsubsection{Ratio of file size and number of lines to execution time}
\label{execution time vs file size vs number lines factor}
\includegraphics{plot-normal-filesize-numlines-factor.ps}

\subsection{Rule ordering}

\subsubsection{Normal vs shuffled vs reversed ordering}
\label{normal vs shuffled vs reversed ordering}
\includegraphics{plot-normal-shuffle-reverse.ps}

\subsubsection{Percentage increase of shuffled and reversed over normal}
\label{normal vs shuffled vs reversed ordering factor}
\includegraphics{plot-normal-shuffle-reverse-factor.ps}

The worst-case performance of shuffled rules can be seen at log file 30,
where it is equivalent to reversed rules.

XXX THIS NEXT PARAGRAPH MIGHT NEED TO BE REWRITTEN\@.

The dips at logfiles 22 and 62--68 correspond to peaks in log file size in
graph~\ref{execution time vs file size vs number lines}.  The explanation
took some time to arrive at, but turns out to be reasonably simple.
Examining the database after parsing logfile 22 shows 64 rules matched
during parsing, but the distribution of matching rules is extremely skewed,
as shown in graph~\ref{log22 distribution by rule}.  Graph~\ref{rule
distribution by program} shows rule distribution by program.  Comparing the
number of log lines per program and rules per program shows that
\textit{postfix/qmgr\/} has the highest number of log lines but has a low
number of rules to try, so the average execution time for a
\textit{postfix/qmgr\/} is quite low, reducing the overall average
execution time per log line.

XXX CREATE A GRAPH AVERAGE EXECUTION TIME PER LOG FILE\@.

\subsubsection{Log 22 distribution by rule}
\label{log22 distribution by rule}
\includegraphics{plot-log22-distribution-by-rule.ps}

The individual rules are not labelled on the x-axis because doing so makes
the graph illegible.

\subsubsection{Log 22 distribution by program}
\label{log22 distribution by program}
\includegraphics{plot-log22-distribution-by-program.ps}

\subsubsection{Rule distribution by program}
\label{rule distribution by program}
\includegraphics{plot-log22-rule-count-by-program.ps}

\subsection{Caching regexs}

\subsubsection{Regex: cached vs discarded}
\label{normal regex vs discard regex}
\includegraphics{plot-cached-discarded.ps}

\subsubsection{Regex: percentage execution time increase}
\label{normal regex vs disarded regex factor}
\includegraphics{plot-cached-discarded-factor.ps}

Two large dips can be seen in graph~\ref{normal regex vs disarded regex
factor} at log files 22 and 62--68, corresponding to the spikes in log file
size in graph~\ref{execution time vs file size vs number lines}.  The
anomaly at files 62--68 is caused by a combination of a mail loop and a
delivery failure which resulted in a huge number of log lines, all of which
were matched by a single rule.  Sorting the rules by the number of
successful matches has the effect that this rule will be tried first, so
the average number of regexs compiled per line drops sharply for those
seven log files, resulting in a much lower increase in relative execution
time, as detailed below.

\begin{tabular}[]{lrr}
                                    & Mean      & Standard Deviation    \\
    All logs                        & 465.58\%  & 114.70\%              \\
    All logs except 22 and 62--68   & 497.43\%  & 50.58\%               \\
    Logs 22 and 62--68              & 127.16\%  & 21.28\%               \\
